\documentclass[10pt]{article}
\voffset=-50pt
\hoffset=-90pt
\textwidth= 500pt
\textheight= 625pt

\newcommand{\myname}{Harjot Gill, Tiernan Garsys, Sam Raper}

\renewcommand{\labelitemi}{$\bullet$}
\renewcommand{\labelitemii}{$\bullet$}
\renewcommand{\labelitemiii}{$\bullet$}
\renewcommand{\labelitemiv}{$\bullet$}

\usepackage{fancyhdr}
\pagestyle{fancy}

\lhead{
  {\bf CIS 700 Airplanes Group 2 Report}\\*
  \myname 
}

\newcommand{\comment}[1]{}

\newcommand{\ms}[1] {
  \texttt{#1}
}

\title{Airplanes Group 2 Report}
\author{\myname}

%%%
\begin{document}
\pagestyle{empty}
\maketitle

%%%
\newpage
\pagestyle{fancy}
\setlength\headheight{60pt}
\setcounter{tocdepth}{3}
\tableofcontents

%%%
\newpage
\section{Introduction}

%%%
\newpage
\section{Initial Insights and Observations}
When we first started working on this project, our initial intuition was to solve the problem 
at a single-agent level without doing any sort of pre-calculation or simulation; one would 
simply need to develop some simple rules for directing the airplanes toward their goal while
avoiding other airplanes, and the rest should fall into place. Our initial implementation, to
this end, was a simple agent wherein each plane would be influenced by a ``force vector'' at each
timestep, similar to a ``boid'' simulation seen in computer graphics. The basis for this force
vector would be a vector pointing from the plane's current location to its destination. At every
time step, one would modify this vector by adding a ``repulsive'' vector pointed away from any
nearby planes or walls. Once the final vector was calculated, the plane would manuever toward
this vector unwaveringly. \\\\
Additionally, we made the insight (along with several other groups) that the total runtime for
any particular map was bounded by the maximum across all planes $P$ of $distance(P_{source}, 
P_{destination}) + P_{departure time}$. In light of this, our initial vector implementation 
contained a prioritization mechanism wherein planes which would land later in an ideal solution
were granted precedence when deciding which plane manuevered out of the way during collision 
scenarios.\\\\
Upon testing our player, it became apparent that our initial approach was insufficient for this
problem. While scenarios with very few planes were solved easily, the scaling of the problem to
maps with tens of airplanes would result in frequent collisions as planes attempts to avoid one
another would inevitably result in collisions with other planes within the simulation. This 
limitation primarily arose from the fact that each agent only considered the immediate state
of the board at each timestep, disregarding both the implications of its potential move and
the potential actions of all other planes in the simulation; by disregarding such information,
it was easy to fall into cases where the ``optimal'' actions taken by any two individual planes 
in the simulation would lead them to a course which was uncorrectable, and thus would result in
a collision. From this, we realized that a more unified, intelligent strategy was necessary to
succeed.

%%%
\newpage
\section{Strategies \& Concepts}
In a radical shift from our initial strategy, we ultimately decided on a strategy that involved
the pre-calculation of flight paths prior to launch. The various aspects of this strategy are 
outlined below.

\subsection{Launch-Time Simulation and Pathfinding}
Instead of attempting to calculate the flight path of a plane dynamically in the air, our final
solution instead set the flight path of a plane once and let it run its course. Once a plane $P_i$'s 
(the i'th plane to depart in the current session)  departure time has been reached in the session, 
the player will begin simulating this plane's path to the destination as follows...
\begin{itemize}
  \item Simulate a path between $P_i$'s source and destination, considering the presence of 
    $P_0$...$P_{i-1}$'s pre-calculated paths with a set of obstacles. On the first iteration, 
    this will be a straight-line path between $P_i$'s source and destination, as no obstacles
    will have been recorded.
  \item If $P_i$ reaches its destination successfully, then set that as the path for $P_i$ in this
    session and launch it. $P_i$ will follow this path until the end of the session.
  \item If $P_i$ experiences a collision during the simulation, then restart the simulation. On this new
    simulation, add an obstacle for the pathfinding where the collision took place.
\end{itemize}

\subsection{Flow Detection}
In response to the trend of ``flow'' boards that arose during this project, our team added a special
``flow detection'' routine during the training phase of the player. If, during training, the player noted
the existence of a ``flow'' on the board (a sequence of five or more planes which approximately shared 
their source, destination, and departure time), then the simulator would, instead of generating a path
as outline above, generate a serialized path where planes would be dispatched in a single-file line
from the flow source to the flow destination.

%%%
\newpage
\section{Implementation}

\subsection{Flight Plan Determination}
At each call of the \ms{updatePlanes()} method, our solution will iterate through the collection of 
planes and perform the appropriate update action based on its current status within the session.
\begin{itemize}
  \item For each plane that has not taken off, we calculate a flight plan using the pathfinding
    implementation described below, which ultimately returns a \ms{List<Waypoint>} representing the waypoints that must
    be traversed by the plane on the way to its destination. The plane is then dispatched heading toward
    its first waypoint.
  \item For each plane that has taken off, check its current bearing and location relative to its
    current waypoint. If it is within a certain radius of its current sought \ms{Waypoint}, pop that \ms{Waypoint} off
    the front of the list and take the next \ms{Waypoint} in the list to be the current \ms{Waypoint}. If the
    plane's bearing is directed toward its current \ms{Waypoint}, then maintain course; else, turn the plane either
    \ms{TURN\_RADIUS} or $Angle(Current Bearing, Goal Bearing)$ degrees (whichever is smaller) toward the current
    waypoint. 
\end{itemize}
    
\subsection{Pathfinding Implementation}
Pathfinding in our implementation is based off of the A* algorithm, using straight-line distance as a
heuristic. Because this heuristic is admissible (i.e. it never over-estimates the actual distance needed
to reach the goal), one knows that for any particular board configuration it will generate the optimal
path between the source and destination. The procedure for determining a path is as follows.
\begin{itemize}
  \item TODO: Make this detailed.
  \item Set some waypoints
  \item Find the shortest path among visible waypoints.
  \item Simulate that path. If it finishes, great! If it doesn't, add an unpathable obstacle around the 
    collision point, and add waypoints around this obstacle. Repeat.
\end{itemize}
In the initial implementation, a collision obstacle resulting from a collision in a prior simulation would
be present in each time step of subsequent simulations, thus causing all pathing decisions to attempt to
move around it. In subsequent implementation, we implemented that the obstacle would only appear in timesteps
around that in which the collision generating the obstacle occurred, to simulate the presence of the collided
plane at that particular point in time.\\\\

\subsection{Referenced Constants}
\begin{itemize}
  \item \ms{TURN\_RADIUS}: The number of degrees the plane may turn during a timestep. The theoretical limit of 
    10.0 by the simulator was found to cause issues due to floating point imperfections, so a value of 
    9.5 was ultimately used.
\end{itemize}


%%%
\newpage
\section{Results}

%%%
\newpage
\section{Contributions}

%%%
\newpage
\section{Future Directions \& Limitations}

\subsection{Flow Optimization}

One minor shortcoming of our solution as presented was its performance 
on so-called ``flow'' boards, characterized by large numbers  of planes that shared 
their source, destination, and departure times and so-named from the serialized 
flow of planes that would form between the source and destination.  While we
were able to improve our performance on these boards by adding flow detection
to the pre-simulation training in the code (implemented by detecting the presence
of five or more planes sharing a source an destination), our solution was limited
by the fact that only one flow of planes was allow between any source-destination
pair. On boards such as \ms{DiagonalFlows} with large amounts of free airspace, Group 5's
player was able to detect the possibility of multiple flows between the source and destination
and subsequently schedule planes to proceed to the destination in two slightly-staggered
flows. While the staggering (necessary for any particular plane to avoid collision with
a plane in another flow immediately at takeoff, before that other plane had cleared the 
airspace) severely reduced the runtime improvements of this strategy, it was nonetheless
better than a one-flow solution; Group 5's player demonstrated a runtime of 666 steps
on \ms{DiagonalFlows}, while our player demonstrated a runtime of 711 steps.\\\\
One could improve on this limitation by adding detection for multiple flow paths between
a source-destination pair during the training phase of our player. Our current implementation
of flow-detection works by finding a shortest path between the source-destination pair, treating
other flows as obstacles obstructing this path. One possibility would be to generate some number
of paths between the source-destination pair, determine which paths are close enough to the
optimal path as to not increase the overall runtime of the simulation after necessary 
staggering was taken into account, and dispatch planes to each of these flows in turn.
Potential implementation difficulties would be being able to determine prior to simulation
that such a splitting would not simply increase the runtime of the entire simulation.

\subsection{Pathfinding Prioritization / Sorting}

Another problem with our solution was the possibility of giving planes whose paths were 
determined last in the sequence of planes overly long paths. As outlined above, our
method of determining paths was greedy in that we would determine the path for any 
particular plane $P_i$ by simply simulating the shortest A* path between the source
and destination of $P_i$ in an environment with planes $P_0$...$P_{i-1}$, resetting
the simulation and trying again with a obstacle placed at the collision point in the
event of a collision. This methodology resulted in a greater number of collisions for the
last planes to have their path decided, which would lead them to be given longer paths to
avoid collisions. Problems arose in that the ordering for resolving plane paths was more-or-less
arbitrary; it was very possible that a plane with a short path in an optimal solution would be
given a longer path, potentially to the detriment of simulation runtime, due to the fact
that it had to consider more obstacles than other planes in determining its final path.\\\\
We attempted to address this problem by prioritizing the order with which planes' paths 
were determined in our pre-flight simulations. Methods tried include...
\begin{itemize}
  \item \ms{Shortest Path First:} Order the planes in ascending order by path length, and 
    resolve flight paths in that order. The intuition behind this was that shorter paths
    would have fewer intersections with other paths, and thus their resolution would generate
    fewer obstacles for later-resolved flights.
  \item \ms{Longest Path First:} Order the planes in descending order by path length, and
    resolve flight paths in that order. The intuition behind this was that longer flights
    are more likely to be the limiting factor in the overall runtime of the simulation, 
    so resolving them first would ensure their runtime would not be increased by collisions
    with shorter flights.
  \item \ms{Least Intersections First:} Determine the straight line paths between all
    source-destination pairs in the simulation. Order the planes in descending order by
    number of intersections with other straight line paths, and resolve flight paths in 
    that order. This method attempted to resolve flights that would be interfered with 
    by many other flights first, thus prevent their runtimes from skyrocketing.
\end{itemize}
In experiments, we found that each of the above methods yielded superior results in different
simulations, with no clear trend of certain strategies working better on certain maps. Due to the
fact that our implementation of sorting was incompatible with our flow detection, our final solution
ultimately scrapped prioritization of plane flights. One issue that would have to be solved if this
were implemented in the future would be gathering useful information for prioritizing flights from
the information that is available at the beginning of the simulation. One only knows when the simulation
starts what the source, destination, and departure time of each plane is. As of time of writing, we 
were unable to find any way of extrapolating from this data a prioritization that would reliably yield
better results on most boards.

%%%
\newpage
\section{Acknowledgments}

\ms{Chris Murphy}, for the awesome class.\\\\
\ms{Tanveer Gill}, for having helped developed the A* package used by our group with Harjot during
the Mosquitos project.

%%%
\newpage
\section{Conclusion}
  
\end{document}
